WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2018-12-22 paper/ms.tex
add comments from all coauthors

\begin{abstract}
    The Transiting Exoplanet Survey Satellite (TESS) recently observed
    18 transits of the hot Jupiter WASP-4b.  The sequence of transits
    occurred $81.6 \pm 11.7$~seconds earlier than had been predicted,
    based on data stretching back to 2007.  This is unlikely
    to be the result of a clock error, because TESS observations of
    other hot Jupiters (WASP-6b, 18b, and 46b) are compatible with a
    constant period, ruling out an 81.6-second offset at the
    6.4$\sigma$ level.  The 1.3-day orbital period of WASP-4b appears
    to be decreasing at a rate of $\dot{P} = -12.6 \pm 1.2$
    milliseconds per year.  This might be caused by tidal orbital
    decay or apsidal precession, although both interpretations have
    shortcomings.  The gravitational influence of a third body is another
    possibility, although at present there is no evidence for such a
    body. Further observations are needed to confirm and
    understand the timing variation.
\end{abstract}

@@

Although the Transiting Exoplanet Survey Satellite (TESS) is designed
to detect new planets, it will also perform precise photometric
monitoring of most of the planets that have been discovered by
ground-based transit surveys over the last two decades.  One of the
applications of the new TESS data is to search for timing anomalies in
hot Jupiter systems.  Long-term monitoring of hot Jupiter transit and
occultation times should eventually reveal two distinct processes.

@@

Second, long-term timing studies may reveal apsidal precession, {\it
  i.e.}, the rotation of the orbital ellipse within the orbital
plane. Apsidal precession has long been observed in eclipsing binaries
\citep[{\it e.g.},][]{
  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
For exoplanets, precessing orbits have been detected in a handful of
cases, including circumbinary systems (Kepler-413
\citealt{kostov_kepler-413b_2014}; Kepler-453
\citealt{welsh_kepler_2015}, planets orbiting rapidly rotating stars
(Kepler-13
\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015};
and systems with multiple interacting giant planets (Kepler-108
\citealt{mills_kepler-108_2017}; Kepler-693
\citealt{masuda_eccentric_2017}).  Apsidal precession is not regularly
observed in hot Jupiter systems because their orbits are nearly
circular, which may itself be a consequence of star-planet tides.  If
it could be observed, then \citet{ragozzine_probing_2009} showed that
it would be possible to measure the planet's Love number, which would
be a valuable constraint on the planet's interior.

So far, the most convincing direct evidence for either orbital decay
or apsidal precession has come from WASP-12b.
\citet{maciejewski_departure_2016} and \citet{patra_2017} showed that
the transit period of WASP-12b has been steadily decreasing by about
30~milliseconds per year over the last decade.  Here, we present
evidence for a timing anomaly in the WASP-4 system.  The hot Jupiter
WASP-4b orbits a G7V star every 1.34 days, corresponding to an orbital
distance of 5.5 stellar radii
\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  It is a good target to
search for departures from a constant period, because transit data
exists all the way back to 2007.  The orbital eccentricity is less
than 0.018 (2$\sigma$), based on the work of
\cite{knutson_friends_2014}, who combined the available transit times,
occultation times, and Doppler data.  The stellar obliquity (or at
least its sky projection) is also compatible with zero, within about
10 degrees
\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.

In what follows, \S~\ref{sec:observations} presents the new TESS
observations, and \S~\ref{sec:timing} describes our timing
analysis.  We tried fitting the data with three models: a constant
period, a decaying period, and a slightly eccentric precessing orbit.
A constant period can be ruled out, but we cannot distinguish between
the latter two possibilities.  Either possibility would have
interesting implications, described in \S~\ref{sec:implications}.
Appendix~\ref{sec:verify_tess} considers the possibility that the
WASP-4 timing anomaly is due to an error in time stamps in the TESS
data products.  We found this possibility to be unlikely because none
of the other hot Jupiters we examined show a timing offset with the
same amplitude as was seen for WASP-4.

@@

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
      {\bf TESS observations of WASP-4b.} On the left, black points
      are TESS flux measurements, with a vertical offset applied. Blue
      curves are best-fit models. The numbers printed next to each
      lightcurve are the approximate transit times expressed in
      BJD minus 2{,}457{,}000.  The right side shows the
      residuals.
       \label{fig:lightcurves}
    }
\end{figure}

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{f2.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Phase-folded lightcurve of WASP-4b.} Black points are
        TESS data. Yellow points are binned measurements.  The bottom
        panel shows the residuals.  The fit to the phase-folded
        transit (blue line) is used when measuring mid-transit times for
        the individual transits (see text).
        \label{fig:phasefold}
    }
\end{figure}

@@

WASP-4 was observed by TESS with Camera 2 from August 23 to September
20, 2018, within the second ``sector'' of science operations.  The
star is designated as TIC 402026209 in the TESS Input Catalog
\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
surrounding WASP-4 were averaged into 2-minute stacks by the onboard
computer.  The data were downlinked via the Deep Space Network, and the
spacecraft time stamps were calibrated against the ground-station clocks.
The spacecraft clock times were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
the known spacecraft trajectory to compute the barycentric time
corrections on a target by target basis, and expressed the timestamps
as Barycentric TESS Julian Dates (BTJD), which is simply the
barycentric Julian date minus 2{,}457{,}000.  The lightcurves were
vetted and released by the MIT TESS Science Office to the Mikulski
Archive for Space Telescopes on November 29,
2018~\citep{ricker_tess_alerts_2018}.

We began our analysis with the Presearch Data Conditioning (PDC)
lightcurve \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}.
We processed the lightcurve as follows.  First, we removed all points
with non-zero quality flags.  This removed data contaminated by coarse
spacecraft pointing, cosmic rays, and other annoyances
\citep{tess_data_product_description_2018}.  We also removed the data
within the first and last hours of both orbits, because of ramp-like
systematic effects that appear during those time ranges. We also
removed the data that might have been adversely affected by ``momentum
dumps,'' the firing of thrusters and resetting of reaction wheels\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}}
that took place every 2.5 days during sector 2.
The data during these events were assigned quality flags corresponding
to ``Reaction Wheel Desaturation Event'' and ``Manual Exclude''.  For
WASP-4, these flags were simultaneously set for 54 distinct cadences,
and there were 10 momentum dumps, averaging about 10 minutes of
flagged data per dump.  Out of caution, we clipped out an additional
10 minutes before and after every momentum dump.

All told, we removed 8\% of the original data points, and were left
with 18{,}165 measurements of the relative flux of WASP-4.  We
normalized the data by dividing out the median flux.  We converted the
timestamps from BTJD into BJD by adding the appropriate 2457000 day
offset \citep{tess_data_product_description_2018}.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
``flatten'' the lightcurves, as is often done with splines,
polynomials, or Gaussian processes.  Instead, we modeled the
out-of-transit flux variations simultaneously with the transit parameters,
as described below.

@@

Using the cleaned PDC lightcurve, we used the Box Least Squares
algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
period, transit duration, and a reference epoch using the TESS data
alone.  Based on the results, we isolated the data within 4 transit
durations of each transit midpoint.  To find the transit parameters
that best fit the data, we first fitted a line to the out-of-transit
flux measurements surrounding each transit, and divided it out.  We
then created a phase-folded lightcurve and fitted a standard transit
model using the analytic formulae given by
\citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.  We assumed the orbit to be circular.
The free parameters were the reference epoch, the planet to star
radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
radius ratio $a/R_\star$, the inclination $i$, two quadratic
limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
orbital period $P$.

@@

To measure the transit times, we returned to the `cleaned' PDC time
series and fitted the data surrounding each transit separately, with
four free parameters: the time of mid-transit $t_{\rm tra}$, the
planet-to-star radius ratio, and the slope and intercept of a linear
trend to account for any slow variations unrelated to the transit.  We
fixed the remaining parameters at the values that had been determined
from the phase-folded TESS lightcurve.  The uncertainty in each data
point was set equal to the root-mean-square (rms) level of the
out-of-transit data.

To verify that the measured uncertainties are plausible, we computed
the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
transit times.  We found that $\chi^2 = 9.2$, with $n=16$ degrees of
freedom.  The variance of the $\chi^2$ distribution is $2n$, so we
would expect $\chi^2 = 16 \pm 5.7$.  Visually inspecting the residuals
showed that the error variance had been overestimated, so we
multiplied the measured TESS errors by a factor $f=0.76$, forcing a
reduced $\chi^2$ of unity.  This lowered the mean uncertainty of the
transit midtimes from $29.8$ to $22.6$ seconds.  We verified that
omitting this step did not appreciably alter any of our conclusions.

Figure~\ref{fig:lightcurves} shows the lightcurve of each transit, the
best-fit models, and the residuals.  Table~2 reports the mid-transit
times and their uncertainties.  After binning the residuals to 1-hour
windows, the lightcurves have an rms scatter of 586\,{\rm ppm}.  The
pre-launch TESS noise model\footnote{\url{github.com/lgbouma/tnm}} \citep{winn_photonflux_2013,Sullivan_2015}
would have predicted an error
budget consisting of the following terms added in quadrature:
410\,{\rm ppm} from photon-counting noise, 202\,{\rm ppm} from
detector read noise, and 673\,{\rm ppm} from the zodiacal background
light.  The level of background light appears to have been
overestimated.

@@

\startdata
{\it Transit/RV parameters:} & & & \\
  $R_{\rm p}/R_\star$                        & $0.15201$              & $+0.00040$, $-0.00033$      & A \\
  $i$~[deg]                                  & $89.06$                & $+0.65$, $-0.84$            & A \\
  $a/R_\star$                                & $5.451$                & $+0.023$, $-0.052$          & A \\
  $u_{\rm linear}$                           & $0.382$                & ---                         & A \\
  $u_{\rm quad}$                             & $0.210$                & ---                         & A \\
  $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
{\it Stellar parameters:} & & & \\
  $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & C \\
  $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & C \\
  $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & C \\
  $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
  $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
  $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
  $\rho_\star$~[g~cm$^{-3}$]                 & $1.711$                & $+0.022$, $-0.048$          & E \\
  $M_\star$~[M$_{\odot}$]                    & $0.864$                & $+0.084$, $-0.090$          & E \\
  $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
{\it Planetary parameters:} & & & \\
  $a$~[AU]                                   & $0.0226$               & $+0.0007$, $-0.0008$        & E \\
  $M_p$~[M$_{\rm Jup}$]                      & $1.186$                & $+0.090$, $-0.098$          & E \\
  $R_p$~[R$_{\rm Jup}$]                      & $1.321$                & $\pm 0.039$                 & E \\
\enddata

\tablecomments{
  (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
  Orbital periods are in Table~4. The limb darkening parameters were
  allowed to float around the \citet{claret_limb_2017} prediction, but
  were unconstrained.
  (B) \citet{triaud_spin-orbit_2010}.
  (C) From HARPS spectra \citep{doyle_accurate_2013}.
  (D) \citet{stassun_accurate_2017}.
  (E) From functions of stellar, transit or RV parameters as discussed in text.
  (F) \citet{gaia_collaboration_gaia_2018}.
  (G) \citet{stassun_TIC_2018}.
}

\end{deluxetable}

@@

Table~1 gives the resulting parameters, which we adopted for the
remaining analysis.  Our parameters are in agreement with those of
previous investigators
\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
By comparing the star's luminosity and spectroscopic parameters with
the outputs of the Yonsei-Yale stellar-evolutionary models, we found
that WASP-4 is a main-sequence star with an age of approximately
$7\,{\rm Gyr}$.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
		The residuals are the observed times minus the calculated times
		assuming a constant period.  The darkest points correspond to
		the most precise data. The constant-period model (gray line) is
		a poor description of the data.  Models with a uniformly
		decreasing period (blue) or apsidal precession (orange) provide
		better fits to the data.  The red square represents the
		combination of all the TESS data and is for display purposes
		only.  The models were fitted to all of the individual transit
		times.
		\label{fig:times}
	}
\end{figure*}

@@

The earliest epoch is from EulerCam on the 1.2-m Euler telescope
\citep{wilson_wasp-4b_2008}.  The second epoch is based on $z$-band
photometry acquired by \citet{gillon_improved_2009} at the VLT 8.2-m
with FORS2.  Subsequent observations were performed by
\citet{winn_transit_2009}, \citet{dragomir_terms_2011},
\citet{sanchis-ojeda_starspots_2011}, \citet{nikolov_wasp-4b_2012},
\citet{hoyer_tramos_2013}, and \citet{ranjan_atmospheric_2014}.
Finally, \citet{huitson_gemini_2017} acquired optical transit spectra
with the 8.1-m Gemini South telescope between 2011 and 2014, one
transit per season.  The per-point standard deviation of their
lightcurves was a few hundred parts per million.  The average
precision in their reported transit times was 5.6~seconds.  Since
those data points carry significant weight in the analysis, we
consulted with those authors to confirm that the time stamps in their
data represent mid-exposure times, that the barycentric correction was
performed correctly, and that the time system of the final results was
BJD$_{\rm TDB}$.  These same authors also used the same instrument
(and largely the same software) to analyze other hot Jupiters, none of
which showed a departure from a constant-period model.

We also compiled all the available occultation times, which are given
in Table~3.  The tabulated values have been corrected for the
light-travel time across the diameter of the orbit by subtracting
$2a/c = 22.8$ seconds from the observed time.
\citet{beerer_secondary_2011} observed two occultations of WASP-4b
using warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
\citet{caceres_ground-based_2011} detected an occultation from the
ground in the $K_S$ band, and gave a time in HJD, without specifying
the time standard.  We assumed the standard was UTC, and performed the
appropriate corrections to convert to BJD$_{\rm TDB}$.  We
also verified that none of our conclusions would be changed if this
assumption was mistaken. Finally, \citet{zhou_secondary_2015} observed an
occultation with the Anglo-Australian Telescope.  They did not
report the observed midpoint, but they did report a result for
$e\cos\omega$ based upon the observed midpoint.  We calculated the
implied midpoint using the formula \citep[{\it
e.g.},][]{winn_exoplanet_2010}
\begin{equation}
  t_{\rm occ}(E) =
  t_0 +  P E  +
  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right).
  \label{eq:occultation_time}
\end{equation}
In total, there are four occultation times.

@@

First, we fitted a constant-period model (``linear ephemeris'') to the
pre-TESS data, and used it to extrapolate to the epochs of the TESS
observations.  The residuals of the best fitting model are shown in
Figure~\ref{fig:arrived_early}.  The transits observed by TESS
occurred earlier than expected.  Because the TESS mission is still in
an early stage, we were concerned about a possible offset in the TESS
time stamps due to an error with the TESS clock or the data processing
pipeline.  Appendix~\ref{sec:verify_tess} describes some tests that
convinced us that a simple offset is unlikely. Assuming that the
observed timing variation is astrophysical, we proceeded by exploring
three models for the timing data in a manner identical to the study
by~\citet{patra_2017}.

>>>>

\begin{abstract}
    The Transiting Exoplanet Survey Satellite (TESS) recently observed
    18 transits of the hot Jupiter WASP-4b.  The sequence of transits
    occurred $81.6 \pm 11.7$~seconds earlier than had been predicted,
    based on data stretching back to 2007.  This is unlikely to be the
    result of a clock error, because TESS observations of other hot
    Jupiters (WASP-6b, 18b, and 46b) are compatible with a constant
    period, ruling out an 81.6-second offset at the 6.4$\sigma$ level.
    The 1.3-day orbital period of WASP-4b appears to be decreasing at
    a rate of $\dot{P} = -12.6 \pm 1.2$ milliseconds per year.  The
    apparent period change might be caused by tidal orbital decay or
    apsidal precession, although both interpretations have
    shortcomings.  The gravitational influence of a third body is
    another possibility, although at present there is no evidence for
    such a body. Further observations are needed to confirm and
    understand the timing variation.
\end{abstract}

@@

Although the Transiting Exoplanet Survey Satellite (TESS) is designed
to detect new planets, it is also performing precise photometric
monitoring of most of the planets that have been discovered by
ground-based transit surveys over the last two decades.  One of the
applications of the new TESS data is to search for timing anomalies in
hot Jupiter systems.  Long-term monitoring of hot Jupiter transit and
occultation times should eventually reveal two distinct processes.

@@

Second, long-term timing studies may reveal apsidal precession, {\it
i.e.}, the rotation of the orbital ellipse within the orbital plane.
Apsidal precession has long been observed in eclipsing binaries
\citep[{\it e.g.},][]{
  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
For exoplanets, precessing orbits have been detected in a handful of
cases, including circumbinary systems (Kepler-413
\citealt{kostov_kepler-413b_2014}; Kepler-453
\citealt{welsh_kepler_2015}), planets orbiting rapidly rotating stars
(Kepler-13
\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015});
and systems with multiple interacting giant planets (Kepler-108
\citealt{mills_kepler-108_2017}; Kepler-693
\citealt{masuda_eccentric_2017}).  Apsidal precession is not regularly
observed in hot Jupiter systems because their orbits are nearly
circular, which may itself be a consequence of star-planet tides.  If
it could be observed, then as \citet{ragozzine_probing_2009} showed it
would be possible to measure the planet's Love number, which would be
a valuable constraint on the planet's interior.

So far, the most convincing direct evidence for either orbital decay
or apsidal precession has come from WASP-12b.
\citet{maciejewski_departure_2016} and \citet{patra_2017} showed that
the transit period of WASP-12b has been steadily decreasing by about
30~milliseconds per year over the last decade.  Here, we present
evidence for a timing anomaly in the WASP-4 system.  The hot Jupiter
WASP-4b orbits a G7V star every 1.34 days, corresponding to an orbital
distance of 5.5 stellar radii
\citep{wilson_wasp-4b_2008,huitson_gemini_2017}.  It is a good target
to search for departures from a constant period, because transit data
exist all the way back to 2007.  The orbital eccentricity is less than
0.018 (2$\sigma$), based on the work of \cite{knutson_friends_2014},
who combined the available transit times, occultation times, and
Doppler data.  The sky projection of the stellar obliquity
is also compatible with zero, within about 10 degrees
\citep{triaud_spin-orbit_2010,sanchis-ojeda_starspots_2011}.

In what follows, \S~\ref{sec:observations} presents the new TESS
observations, and \S~\ref{sec:timing} describes our timing analysis.
We tried fitting the data with three models: a constant period, a
decaying period, and a slightly eccentric precessing orbit.  A
constant period can be ruled out, but we cannot distinguish between
the latter two possibilities.  Either possibility would have
interesting implications, described in \S~\ref{sec:implications}.
Appendix~\ref{sec:verify_tess} considers the possibility that the
WASP-4 timing anomaly is due to an error in timestamps in the TESS
data products.  We found this possibility to be unlikely because none
of the other hot Jupiters we examined show a timing offset with the
same amplitude as was seen for WASP-4.

@@

\begin{figure*}[t]
    \begin{center}
		\includegraphics[height=0.97\textheight]{f1.png}
    \end{center}
    \vspace{-0.8cm}
    \caption{
      {\bf TESS observations of WASP-4b.} On the left, black points
      are TESS flux measurements, with a vertical offset applied. Blue
      curves are best-fit models. The numbers printed next to each
      lightcurve are the approximate transit times expressed in BJD
      minus 2{,}457{,}000.  The right side shows the residuals.
       \label{fig:lightcurves}
    }
\end{figure*}

\begin{figure*}[t]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{f2.png}
    \end{center}
    \vspace{-0.7cm}
    \caption{
        {\bf Phase-folded lightcurve of WASP-4b.} Black points are
        TESS data. Yellow points are binned measurements.  The bottom
        panel shows the residuals.  The fit to the phase-folded
        transit (blue line) is used when measuring mid-transit times
        for the individual transits (see text).
        \label{fig:phasefold}
    }
\end{figure*}

@@

WASP-4 was observed by TESS with Camera 2 from August 23 to September
20, 2018, within the second ``sector'' of science operations.  The
star is designated as TIC 402026209 in the TESS Input Catalog
\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
surrounding WASP-4 were averaged into 2-minute stacks by the onboard
computer.  The data were downlinked via the Deep Space Network, and
the spacecraft timestamps were calibrated against the ground-station
clocks.  The spacecraft clock times were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
the known spacecraft trajectory to compute the barycentric time
corrections on a target by target basis, and expressed the timestamps
as Barycentric TESS Julian Dates (BTJD), which is simply the
barycentric Julian date minus 2{,}457{,}000.  Lightcurves that were
flagged as Threshold Crossing Events by the SPOC pipeline were then
vetted and released by the MIT TESS Science Office to the Mikulski
Archive for Space Telescopes on November 29,
2018~\citep{ricker_tess_alerts_2018}.

We began our analysis with the Presearch Data Conditioning (PDC)
lightcurve \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}.
We processed the lightcurve as follows.  First, we removed all points
with non-zero quality flags.  This removed data that might have been
adversely affected by ``momentum dumps,'' the firing of thrusters and
resetting of reaction
wheels\footnote{\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
that took place every 2.5 days during sector 2
\citep{tess_data_product_description_2018}.  The data during these
events were assigned quality flags corresponding to ``Reaction Wheel
Desaturation Event'' and ``Manual Exclude''.  For WASP-4, these flags
were simultaneously set for 54 distinct cadences, and there were 10
momentum dumps, averaging about 10 minutes of flagged data per dump.
Out of caution, we clipped out an additional 10 minutes before and
after every momentum dump.  We also removed the data within the first
and last hours of both orbits, because of ramp-like systematic effects
that appear during those time ranges.

All told, we removed 8\% of the original data points, and were left
with 18{,}165 measurements of the relative flux of WASP-4.  We
normalized the data by dividing out the median flux.  We converted the
timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
appropriate 2457000 day offset
\citep{tess_data_product_description_2018}.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
``flatten'' the lightcurves, as is often done with splines,
polynomials, or Gaussian processes.  Instead, we modeled the
out-of-transit flux variations simultaneously with the transit
parameters, as described below.

@@

Using the cleaned PDC lightcurve, we used the Box Least Squares
algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
period, transit duration, and a reference epoch using the TESS data
alone.  Based on the results, we isolated the data within 4 transit
durations of each transit midpoint.  To find the transit parameters
that best fit the data, we first fitted a line to the out-of-transit
flux measurements surrounding each transit, and divided it out.  We
then created a phase-folded lightcurve and fitted a standard transit
model using the analytic formulae given by
\citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.  We assumed the orbit to be circular,
consistent with the limits from radial velocities and occultation
timing
\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
The free parameters were the reference epoch, the planet to star
radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
radius ratio $a/R_\star$, the inclination $i$, two quadratic
limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
orbital period $P$.

@@

To measure the transit times, we returned to the `cleaned' PDC time
series and fitted the data within four transit durations of each
transit separately. We used four free parameters: the time of
mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
slope and intercept of a linear trend to account for any slow
variations unrelated to the transit.  We fixed the remaining
parameters at the values that had been determined from the
phase-folded TESS lightcurve.  The uncertainty in each data point was
set equal to the root-mean-square (rms) level of the out-of-transit
data.

To verify that the measured uncertainties are estimated accurately, we
computed the reduced $\chi^2$ for a linear ephemeris fit to the
measured TESS mid-transit times.  We found that $\chi^2 = 9.2$, with
$n=16$ degrees of freedom.  The variance of the $\chi^2$ distribution
is $2n$, so we would expect $\chi^2 = 16 \pm 5.7$.  Visually
inspecting the residuals showed that the error variance had been
overestimated, so we multiplied the measured TESS errors by a factor
$f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
uncertainty of the transit midtimes from $29.8$ to $22.6$ seconds.  We
verified that omitting this step did not appreciably alter any of our
conclusions.

Figure~\ref{fig:lightcurves} shows the lightcurve of each individual
transit, the best-fit models, and the residuals.  Table~2 reports the
mid-transit times and their uncertainties.  After binning the
residuals to 1-hour windows, the lightcurves have an rms scatter of
586\,{\rm ppm}.  The pre-launch TESS noise
model\footnote{\url{github.com/lgbouma/tnm}}
(\citealt{winn_photonflux_2013}, \citealt{Sullivan_2015} Section 6.4)
would have predicted an error budget consisting of the following terms
added in quadrature: 410\,{\rm ppm} from photon-counting noise,
202\,{\rm ppm} from detector read noise, and 673\,{\rm ppm} from the
zodiacal background light.  The level of background light appears to
have been overestimated.

@@

\startdata
{\it Transit/RV parameters:} & & & \\
  $R_{\rm p}/R_\star$                        & $0.15201$              & $+0.00040$, $-0.00033$      & A \\
  $i$~[deg]                                  & $89.06$                & $+0.65$, $-0.84$            & A \\
  $a/R_\star$                                & $5.451$                & $+0.023$, $-0.052$          & A \\
  $u_{\rm linear}$                           & $0.382$                & ---                         & A \\
  $u_{\rm quad}$                             & $0.210$                & ---                         & A \\
  $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
{\it Stellar parameters:} & & & \\
  $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & C \\
  $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & C \\
  $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & C \\
  $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
  $A_V$~[mag]                                & $0.03$                 & $+0.02, -0.01$              & D \\
  $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
  $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
  $\rho_\star$~[g~cm$^{-3}$]                 & $1.711$                & $+0.022$, $-0.048$          & E \\
  $M_\star$~[M$_{\odot}$]                    & $0.864$                & $+0.084$, $-0.090$          & E \\
  $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
  %Rotation period~[days]                     & $\approx$34            & ---                         & H \\
{\it Planetary parameters:} & & & \\
  $a$~[AU]                                   & $0.0226$               & $+0.0007$, $-0.0008$        & E \\
  $M_p$~[M$_{\rm Jup}$]                      & $1.186$                & $+0.090$, $-0.098$          & E \\
  $R_p$~[R$_{\rm Jup}$]                      & $1.321$                & $\pm 0.039$                 & E \\
\enddata

\tablecomments{
  (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
  Orbital periods are in Table~4. The limb darkening parameters were
  allowed to float around the \citet{claret_limb_2017} prediction, but
  were unconstrained.
  (B) \citet{triaud_spin-orbit_2010}.
  (C) From HARPS spectra \citep{doyle_accurate_2013}.
  (D) \citet{stassun_accurate_2017}.
  (E) This work, see \S~\ref{sec:system_parameters}.
  (F) \citet{gaia_collaboration_gaia_2018}.
  (G) \citet{stassun_TIC_2018}.
  %(H) \citet{hoyer_tramos_2013}, from star-spot occultation events.
}

\vspace{-1cm}
\end{deluxetable}

@@

Table~1 gives the resulting parameters, which we adopted for the
remaining analysis.  Our parameters are in agreement with those of
previous investigators, but have the benefit of incorporating the {\it
Gaia} parallax
\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013,huitson_gemini_2017}.
By comparing the star's luminosity and spectroscopic parameters with
the outputs of the Yonsei-Yale stellar-evolutionary models, we found
that WASP-4 is a main-sequence star with an age of approximately
$7\,{\rm Gyr}$.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
		The residuals are the observed times minus the calculated times
		assuming a constant period.  The darkest points correspond to
		the most precise data. The constant-period model (gray line) is
		a poor description of the data.  Models with a uniformly
		decreasing period (blue) or apsidal precession (orange) provide
		better fits to the data.  The red square represents the
		combination of all the TESS data and is for display purposes
		only.  The models were fitted to all of the individual transit
		times.
		\label{fig:times}
	}
\end{figure*}

@@

The earliest epoch is from EulerCam on the 1.2-m Euler telescope
\citep{wilson_wasp-4b_2008}.  The second epoch is based on $z$-band
photometry acquired by \citet{gillon_improved_2009} at the VLT 8.2-m
with FORS2.  Subsequent observations were performed by
\citet{winn_transit_2009}, \citet{dragomir_terms_2011},
\citet{sanchis-ojeda_starspots_2011}, \citet{nikolov_wasp-4b_2012},
\citet{hoyer_tramos_2013}, and \citet{ranjan_atmospheric_2014}.
Finally, \citet{huitson_gemini_2017} acquired optical transit spectra
with the 8.1-m Gemini South telescope between 2011 and 2014, one
transit per season.  The per-point standard deviation of their
lightcurves was a few hundred parts per million.  The average
precision in their reported transit times was 5.6~seconds.  Since
those data points carry significant weight in the analysis, we checked
that the timestamps in their data represent mid-exposure times, that
the barycentric correction was performed correctly, and that the time
system of the final results was BJD$_{\rm TDB}$.  These same authors
also used the same instrument and method to analyze other hot
Jupiters, none of which showed a departure from a constant-period
model.  Finally, these authors also measured two transit midpoints
using Spitzer (Baxter et al., in prep); these
times agree with the Gemini South results.

We also compiled the available occultation times, which are given in
Table~3.  The tabulated values have been corrected for the
light-travel time across the diameter of the orbit by subtracting
$2a/c = 22.8$ seconds from the observed time.
\citet{beerer_secondary_2011} observed two occultations of WASP-4b
using warm Spitzer in the 3.6\,$\mu$m and 4.5\,$\mu$m bands.
\citet{caceres_ground-based_2011} detected an occultation from the
ground in the $K_S$ band, and gave a time in HJD, without specifying
the time standard.  We assumed the standard was UTC, and performed the
appropriate corrections to convert to BJD$_{\rm TDB}$.  We verified
that none of our conclusions would be changed if this assumption was
mistaken. Finally, \citet{zhou_secondary_2015} observed an occultation
with the Anglo-Australian Telescope.  They did not report the observed
midpoint, but they did report a result for $e\cos\omega$ based upon
the observed midpoint.  We calculated the implied midpoint using the
formula \citep[{\it e.g.},][]{winn_exoplanet_2010}
\begin{equation}
  t_{\rm occ}(E) =
  t_0 +  P E  +
  \frac{P}{2} \left( 1 + \frac{4}{\pi} e\cos\omega \right).
  \label{eq:occultation_time}
\end{equation}
In total, there are four occultation times.

@@

First, we fitted a constant-period model (``linear ephemeris'') to the
pre-TESS data, and used it to extrapolate to the epochs of the TESS
observations.  The residuals of the best fitting model are shown in
Figure~\ref{fig:arrived_early}.  The transits observed by TESS
occurred earlier than expected.  Because the TESS mission is still in
an early stage, we were concerned about a possible offset in the TESS
timestamps due to an error with the TESS clock or the data processing
pipeline.  Appendix~\ref{sec:verify_tess} describes some tests that
convinced us that a simple offset is unlikely. Assuming that the
observed timing variation is astrophysical, we proceeded by exploring
three models for the timing data in a manner identical to the study
by~\citet{patra_2017}.
